# Preface {.unnumbered}

```{r setup_r, echo=FALSE, cache=TRUE, message=FALSE}
# setup book r environment
install.packages(
  'openxlsx2',
  repos = c('https://janmarvin.r-universe.dev',
            'https://cloud.r-project.org')
)
```

This is a work in progress book describing the features of [`openxlsx2`](https://github.com/JanMarvin/openxlsx2/) [@openxlsx2]. Having written a book before, I never imagined to ever write one again and therefore I shall not do it. But still I consider it a nice addition to have something more flexible as our `vignettes`.

The `openxlsx2` book is a comprehensive guide to using the R package openxlsx2 for working with xlsx files. It covers core functionalities such as reading, writing, and editing office open xml (OOXML) spreadsheet files, alongside advanced features like styling worksheets, handling conditional formatting, creating charts, managing pivot tables, and adding data validation. The book also discusses extending the package's functionality and upgrading from the original openxlsx package. It was created in the hopes that it's a useful resource for R users needing spreadsheet file manipulation.

This manual was compiled using:

```{r}
R.version
```

and

```{r}
packageVersion("openxlsx2")
```

Graphics might reflect earlier states and are not constantly updated. If you find any irregularities where our code produces different output than expected, please let us know in the issue tracker at <https://github.com/JanMarvin/openxlsx2/>.

## Additional examples

For many more examples of what `openxlsx2` can do, have a look at the `Show and tell` section of the `openxlsx2` discussion board: <https://github.com/JanMarvin/openxlsx2/discussions/categories/show-and-tell>

## Where to get help

For all things `openxlsx2` consult our discussion board at <https://github.com/JanMarvin/openxlsx2/discussions/categories/q-a>
